---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
map_theme <- list(
  ggplot2::theme_void(),
  ggplot2::theme(panel.grid.major = ggplot2::element_line(colour = "transparent"))
)

# API calls are slow, and repeated API calls puts unneccessary drain on Transitland,
# so we cache calls.
simpleCache <- function(object,key,refresh=FALSE){
  cache_path=file.path(getOption("cache_path"),key)
  if(!refresh & file.exists(cache_path)) readRDS(cache_path) else {
    data=object
    saveRDS(data,file=cache_path)
    data
  }
}
```


This is a quick demo of how to access Transitland data. As an example we are looking at transit service in the City of Vancouver. There are a number of ways to query transit data, we will specify a bounding box based on the geo polygon for Vancouver that we grabbed using the **cancensus** package.

```{r}
library(transitland)
library(tidyverse)
library(sf)
# library(cancensus)
# region <- list_census_regions("CA16") %>% filter(name=="Vancouver",level=="CSD")
# vancouver <- get_census("CA16",regions=as_census_region_list(region),geo_format = "sf",level = "Regions")
vancouver <- read_sf("../data/vancouver.geojson")
bbox=st_bbox(vancouver)
```
# Transit Routes
To start off we pull the route data for the City of Vancouver using the bounding box from our polygon
```{r}
transit_routes <- simpleCache(get_transit_routes(list(bbox=bbox,per_page=1000),get_all = TRUE),"yvr_routes")
```

and plot it.
```{r}
cols <- transit_routes$color %>% unique
route_colours=setNames(paste0("#",cols),cols)
ggplot(vancouver) +
  geom_sf() +
  map_theme +
  geom_sf(data=transit_routes,aes(color=color)) +
  guides(color=FALSE) +
  scale_color_manual(values=route_colours) +
  coord_sf(xlim=c(bbox$xmin,bbox$xmax),ylim=c(bbox$ymin,bbox$ymax))
```




# Transit Stops
Next we query all transit stops in the bounding box and clip them to the City of Vancouver and identify the high frequency B-line and Skytrain stops
```{r}
stops <- simpleCache(get_transit_stops(list(bbox=bbox,per_page=1000),get_all=TRUE),"yvr_transit_stops") %>%
  filter(st_intersects(.,vancouver)>0)

bline_skytrain_routes="099|095|096| Line"
# label the stops by route
stops <- stops %>% 
  mutate(routes=lapply(routes_serving_stop,function(x){fromJSON(x)$route_name %>% unique()})) %>% 
  mutate(routes_string = lapply(routes,function(x){paste(x,collapse=",")})) %>%
  mutate(bline=grepl(bline_skytrain_routes,routes_string))
```

and visualize the data.
```{r}
ggplot(vancouver) +
  geom_sf() +
  map_theme +
  geom_sf(data=stops,aes(color=bline), alpha=0.8,size=1,show.legend = "point") +
  labs(title="Transit Stops in Vancouver",color="B-Line")
```

# Service Frequency
We are interesting in service frequency, so we pull all departures between 9am and 10am and plot the frequencies by stop.
```{r}
params <- list(bbox=bbox,
               per_page=10000,
               origin_departure_between="09:00:00,10:00:00",
               date="2018-03-13")
schedule_stops <- simpleCache(get_transit_schedule_stops(params,get_all = TRUE),"yvr_schedule_9-10") %>% 
  filter(origin_onestop_id %in% stops$onestop_id )

frequency_counts <- schedule_stops %>% group_by(origin_onestop_id) %>% summarize(count=n())

ggplot(frequency_counts %>% group_by(count) %>% summarize(n=n()), aes(x=count)) +
  geom_histogram(stat="bin",fill="steelblue") +
  theme_bw() +
  labs(title="Number of stops by frequency, 9am to 10am",y="Number of stops","Frequency of service per hour",x="Departures per hour")
```



# All on one map
We can join the frequency on the stations and get a geographic overview of transit frequency by stop, binning the frequency for easier readability.
```{r}

plot_data <- stops %>% select(onestop_id) %>% 
  left_join(frequency_counts %>% rename(onestop_id=origin_onestop_id),by="onestop_id")

#breaks <- classInt::classIntervals(plot_data$count)
breaks <- c(0,1,3,6,12,20,30,40,60,120,360,Inf)
labels <- c("Fewer than 1/hour","hourly - every 20 minutes","every 20 minutes - every 10 minutes",
            "every 10 minutes - every 5 minutes",
            "every 5 minutes - every 3 minutes",
            "every 3 minutes - every 2 minutes",
            "every 2 minutes - every 1.5 minutes",
            "every 1.5 minutes - every minute",
            "every minute - 2 per minute",
            "2 per minute - 3 per minute",
            "more than 3 per minute")
process_plot_data<-function(data){
  data %>% mutate(d=cut(count,breaks=breaks, labels=labels)) %>%
    mutate(d=as.character(d)) %>%
    mutate(d=ifelse(is.na(d),"None",d)) %>%
    mutate(d=factor(d,levels=c("None",labels))) %>%
    mutate(size=ifelse(is.na(count),0,sqrt(count)))
}
plot_data <- plot_data %>% process_plot_data
  

color_map <- setNames(c("grey60",viridis::viridis(length(labels))),c("None",labels))

ggplot(vancouver) +
  geom_sf() +
  map_theme +
  geom_sf(data=transit_routes,color="darkgrey",size=0.2) +
  scale_color_manual(values=color_map) +
  geom_sf(data=plot_data,aes(color=d,size=size), alpha=0.8,show.legend = "point") +
  scale_size_continuous(range=c(0.3,2),guide = FALSE) +
  coord_sf(xlim=c(bbox$xmin,bbox$xmax),ylim=c(bbox$ymin,bbox$ymax)) +
  labs(title="Vancouver transit frequency per stop/station 9am-10am",
       color="Departures/minute",caption="MountainMath, Transitland")
```





# Compaing to evening service
To round up the example we want to compare this to transit departures midnight to 1am.
```{r}
params <- list(bbox=bbox,
               per_page=10000,
               origin_departure_between="24:00:00,25:00:00",
               date="2018-03-13")
schedule_stops_late <- simpleCache(get_transit_schedule_stops(params,get_all = TRUE),"yvr_schedule_24-45") %>% 
  filter(origin_onestop_id %in% stops$onestop_id )

frequency_counts_late <- schedule_stops_late %>% group_by(origin_onestop_id) %>% summarize(count=n())

ggplot(frequency_counts_late %>% group_by(count) %>% summarize(n=n()), aes(x=count)) +
  geom_histogram(stat="bin",fill="steelblue") +
  theme_bw() +
  labs(title="Number of stops by frequency, midnight to 1am",y="Number of stops","Frequency of service per hour",x="Departures per hour")
```

As expected, the frequency is significantly lower at night, but still quite decent. We can again plot the geographic distribution.

```{r}

plot_data <- stops %>% 
  select(onestop_id) %>% 
  left_join(frequency_counts_late %>% rename(onestop_id=origin_onestop_id),by="onestop_id") %>%
  process_plot_data

ggplot(vancouver) +
  geom_sf() +
  map_theme +
  geom_sf(data=transit_routes,color="darkgrey",size=0.2) +
  scale_color_manual(values=color_map) +
  geom_sf(data=plot_data,aes(color=d,size=size), alpha=0.8,show.legend = "point") +
  coord_sf(xlim=c(bbox$xmin,bbox$xmax),ylim=c(bbox$ymin,bbox$ymax)) +
  scale_size_continuous(range=c(0.3,2),guide = FALSE) +
  labs(title="Vancouver transit frequency per stop/station midnight-1am",
       color="Departures/minute",caption="MountainMath, Transitland")
```
